---
title: "Using nVennR to generate and explore n-dimensional, quasi-proportional Venn diagrams"
author: "Victor Quesada"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using nVennR to generate and explore n-dimensional, quasi-proportional Venn diagrams}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, F, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.show='hide'
)
library(knitr)
uuid <- function() {
  hex_digits <- c(as.character(0:9), letters[1:6])
  hex_digits <- toupper(hex_digits)
  paste(sample(hex_digits, 8), collapse='')
}

subsuid <- function(regex, strng){
  l <- gregexpr(regex, strng, perl = T)
  for (x in regmatches(strng, l)){ 
    m <- regexpr('([^\\{ \\.\\#]+)', x, perl = T)
    names <- regmatches(x, m)
    gstr = strng
    for (name in names){
      nname <- paste('([^\\d\\w<>]', name, ')', sep="")
      gstr <- gsub(nname, paste('\\1', '_', uuid(), sep=""), gstr, perl = T) 
    }
    return(gstr)
  }
}

knit_print.nVennR = function(x, ...) {
  if (is.null(x$svg)){
    x <- showSVG(x)
  }
  s <- subsuid('[\\.\\#](.+?)\\s*\\{', x$svg)
  s <- subsuid('bl\\d+', s)
  knitr::asis_output(s)
}
# register the method
registerS3method("knit_print", "nVennObj", knit_print.nVennR)
local({
  hook_source <- knitr::knit_hooks$get('source')
  knitr::knit_hooks$set(source = function(x, options) {
    x <- x[!grepl('#noshow$', x)]
    hook_source(x, options)
  })
})
```

nVennR provides an R interface to the [nVenn algorithm](https://academic.oup.com/bioinformatics/article/34/13/2322/4904268). This vignette intends to illustrate three basic uses of nVennR:

* Create Venn and Euler diagrams
* Generate and customize high-quality figures to depict the diagram
* Explore the regions of the diagram

## Create diagrams
There are two ways to create a Venn diagram object (`nVennObj`), which will be referenced as high-level (by providing intersecting lists) and low-level (from scratch).

### High-level

The most common use for a package like nVennR is to depict the relationships between several intersecting lists. The main function for this task is `plotVenn`. The input is a list of vectors (or lists) describing each set. The name of each inner vector will be used for labeling. If inner vectors are not named, labels can be provided as a vector with `sNames`. Empty names will be filled with `GroupN`. Examples:

```{r "plotVenn"}
library(nVennR)
exampledf
sas <- subset(exampledf, SAS == "Y")$Employee
python <- subset(exampledf, Python == "Y")$Employee
rr <- subset(exampledf, R == "Y")$Employee
myV <- plotVenn(list(SAS=sas, PYTHON=python, R=rr), nCycles = 2000)
myV #noshow

```
